\textbf{Empirical evidence for beneficial mutations that are not adaptive innovations in mammals}

### 2023-04-03 manuscript/supp-mat.tex
Updated manuscript, changed colors.
Updated manuscript before NG corrections

\newcommand{\UniDimArray}[1]{\bm{#1}}
\newcommand{\BiDimArray}[1]{\bm{#1}}
\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\Var}{\textrm{Var}}
\newcommand{\der}{\textrm{d}}
\newcommand{\e}{\textrm{e}}
\newcommand{\avg}[1]{\left< #1 \right>} % for average
\newcommand{\Ne}{N_{\textrm{e}}}
\newcommand{\proba}{\mathbb{P}}
\newcommand{\pfix}{\proba_{\textrm{fix}}}
\newcommand{\dn}{d_N}
\newcommand{\ds}{d_S}
\newcommand{\dnds}{\dn / \ds}
\newcommand{\Sphy}{S_{0}}
\newcommand{\Sphyclass}{\mathcal{C}}
\newcommand{\SphyMean}{\overline{\Sphy}}
\newcommand{\divStrongDel}{\Sphy < -3}
\newcommand{\divDel}{-3 < \Sphy < -1}
\newcommand{\divWeakDel}{-1 < \Sphy < 0}
\newcommand{\divWeakAdv}{0 < \Sphy < 1}
\newcommand{\divAdv}{ \Sphy > 1}
\newcommand{\PdivStrongDel}{\proba \left[ \divStrongDel \right]}
\newcommand{\PdivDel}{\proba \left[ \divDel \right]}
\newcommand{\PdivWeakDel}{\proba \left[ \divWeakDel \right]}
\newcommand{\PdivWeakAdv}{\proba \left[ \divWeakAdv \right]}
\newcommand{\PdivAdv}{\proba \left[ \divAdv \right]}
\newcommand{\given}{\mid}
\newcommand{\Spop}{S}
\newcommand{\SpopMean}{\overline{\Spop}}
\newcommand{\polyDel}{\Spop < -1}
\newcommand{\polyNeutral}{-1 < \Spop < 1}
\newcommand{\polyAdv}{ \Spop > 1}
\newcommand{\PpolyDel}{\proba \left[ \polyDel \right]}
\newcommand{\PpolyNeutral}{\proba \left[ \polyNeutral \right]}
\newcommand{\PpolyAdv}{\proba \left[ \polyAdv \right]}

\title{Up to 25\% of beneficial mutations in protein sequences are not adaptive innovations in mammals}

\author{
    \large
    \textbf{T. {Latrille}$^{1}$, J. {Joseph}$^{2}$, N. {Salamin}$^{1}$}\\
    \normalsize
    $^{1}$Université de Lausanne, Lausanne, Switzerland\\
    $^{2}$Université de Lyon, CNRS, LBBE UMR 5558, Villeurbanne, France \\
    \texttt{\href{mailto:thibault.latrille@ens-lyon.org}{thibault.latrille@ens-lyon.org}} \\
}

@@

    \section{Mutation-selection codon models}\label{sec:site-specific-mutation-selection-codon-models}
    In BayesCode (\url{https://github.com/ThibaultLatrille/bayescode}), mutation-selection codon models are obtained by running \textit{mutselomega} for 2000 points of MCMC with the options:
    \begin{scriptsize}
        \begin{verbatim}
 mutselomega ---omegashift 0.0 --ncat 30 -a my_alignment.phy -t my_tree.newick -u 2000 my_genename
        \end{verbatim}
    \end{scriptsize}
    The collection of site-specific fitness profiles ($\UniDimArray{F^{(i)}}, \forall i$) are then obtained by running \textit{readmutselomega}, reading 1000 points of MCMC (first 1000 are considered as burn-in) with the options:
    \begin{scriptsize}
        \begin{verbatim}
 readmutselomega --every 1 --until 2000 --burnin 1000 --ss my_genename
        \end{verbatim}
    \end{scriptsize}
    The gene-specific mutation matrix ($\UniDimArray{\mu}$) is also obtained by running \textit{readmutselomega}, reading 1000 points of MCMC (first 1000 are considered as burn-in) with the options:
    \begin{scriptsize}
        \begin{verbatim}
 readmutselomega --every 1 --until 2000 --burnin 1000 --nuc my_genename
        \end{verbatim}
    \end{scriptsize}

@@

    \subsubsection{$\dnds$ and beneficial back-mutations}
    \begin{center}
        \captionof{table}{$\dnds$ of beneficial back-mutations.}
        \scriptsize
        \begin{longtable*}{|l|l|r|r|r|}
            \toprule
            Population & Species & Tajima $\theta_{\pi}$ & $\mathbb{P}_{div}[\Sphy > 1]$ & $\dn(\Sphy > 1) / \ds$ \\
            \midrule
            \endhead
            \midrule
            \multicolumn{5}{r}{{Continued on next page}} \\
            \midrule
            \endfoot

@@

            Equus c. &      Equus caballus &               $ 0.002$ &                $ 0.116$ &              $ 1.214$ \\
            Iran &          Bos taurus &               $ 0.005$ &                $ 0.128$ &              $ 1.309$ \\
            Uganda &          Bos taurus &               $ 0.005$ &                $ 0.130$ &              $ 1.352$ \\
            Australia &        Capra hircus &               $ 0.003$ &                $ 0.121$ &              $ 1.147$ \\
            France &        Capra hircus &               $ 0.003$ &                $ 0.121$ &              $ 1.154$ \\
            Iran (C. aegagrus) &        Capra hircus &               $ 0.003$ &                $ 0.121$ &              $ 1.154$ \\
            Iran &        Capra hircus &               $ 0.004$ &                $ 0.122$ &              $ 1.164$ \\
            Italy &        Capra hircus &               $ 0.003$ &                $ 0.120$ &              $ 1.148$ \\
            Morocco &        Capra hircus &               $ 0.004$ &                $ 0.123$ &              $ 1.170$ \\
            Iran &          Ovis aries &               $ 0.007$ &                $ 0.113$ &              $ 1.210$ \\
            Iran (O. orientalis) &          Ovis aries &               $ 0.009$ &                $ 0.112$ &              $ 1.214$ \\
            Iran (O. vignei) &          Ovis aries &               $ 0.007$ &                $ 0.113$ &              $ 1.200$ \\
            Various &          Ovis aries &               $ 0.009$ &                $ 0.112$ &              $ 1.207$ \\
            Morocco &          Ovis aries &               $ 0.007$ &                $ 0.113$ &              $ 1.215$ \\
            Barbados & Chlorocebus sabaeus &               $ 0.004$ &                $ 0.122$ &              $ 1.477$ \\
            Central Afr. Rep. & Chlorocebus sabaeus &               $ 0.006$ &                $ 0.123$ &              $ 1.501$ \\
            Ethiopia & Chlorocebus sabaeus &               $ 0.005$ &                $ 0.123$ &              $ 1.494$ \\
            Gambia & Chlorocebus sabaeus &               $ 0.005$ &                $ 0.123$ &              $ 1.491$ \\
            Kenya & Chlorocebus sabaeus &               $ 0.005$ &                $ 0.123$ &              $ 1.503$ \\
            Nevis & Chlorocebus sabaeus &               $ 0.003$ &                $ 0.123$ &              $ 1.484$ \\
            South Africa & Chlorocebus sabaeus &               $ 0.006$ &                $ 0.125$ &              $ 1.528$ \\
            Saint Kitts & Chlorocebus sabaeus &               $ 0.004$ &                $ 0.123$ &              $ 1.495$ \\
            Zambia & Chlorocebus sabaeus &               $ 0.006$ &                $ 0.123$ &              $ 1.493$ \\
            African &        Homo sapiens &               $ 0.003$ &                $ 0.102$ &              $ 1.676$ \\
            Admixed American &        Homo sapiens &               $ 0.002$ &                $ 0.102$ &              $ 1.669$ \\
            East Asian &        Homo sapiens &               $ 0.002$ &                $ 0.102$ &              $ 1.682$ \\
            European &        Homo sapiens &               $ 0.002$ &                $ 0.101$ &              $ 1.659$ \\
            South Asian &        Homo sapiens &               $ 0.002$ &                $ 0.102$ &              $ 1.676$ \\

@@

    \subsection{Correlation with diversity}

    \subsubsection{Proportion of nearly-neutral mutations ($\polyNeutral$)}\label{subsec:proportion-nearly-neutral-mutations}
    \captionof{figure}{~}
    \begin{minipage}{0.32\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/5bins-all/regression-MutSel/results.watterson.all_P-Seq0.scatter.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/5bins-all/regression-MutSel/results.watterson.neg-weak_P-Seq0.scatter.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/5bins-all/regression-MutSel/results.watterson.pos-weak_P-Seq0.scatter.pdf}
    \end{minipage}
    \begin{itemize}
        \item $\ \proba [ \polyNeutral ]$ is the proportion of DFE with a selection coefficient at the population-genetic scale ($\Spop$) between -1 and 1, estimated for the whole genome.
        \item $\ \proba [ \polyNeutral \given \divWeakDel]$ is the proportion of DFE with a selection coefficient at the population-genetic scale ($\Spop$) between -1 and 1, estimated of SNPs observed with a selection coefficient at the phylogenetic scale ($\Sphy$) between -1 and 0.
        \item $\ \proba [ \polyNeutral \given \divWeakAdv ]$ is the proportion of DFE with a selection coefficient at the population-genetic scale ($\Spop$) between -1 and 1, estimated of SNPs observed with a selection coefficient at the phylogenetic scale ($\Sphy$) between 0 and 1.
    \end{itemize}

    \subsubsection{Proportion of deleterious mutations ($\polyDel$)}\label{subsec:proportion-deleterious-mutations}
    \captionof{figure}{~}
    \begin{minipage}{0.32\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/5bins-all/regression-MutSel/results.watterson.all_P-Sinf0.scatter.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/5bins-all/regression-MutSel/results.watterson.neg_P-Sinf0.scatter.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/5bins-all/regression-MutSel/results.watterson.neg-strong_P-Sinf0.scatter.pdf}
    \end{minipage}
    \begin{itemize}
        \item $\ \proba [ \polyDel ]$ is the proportion of DFE with a selection coefficient at the population-genetic scale ($\Spop$) lower than -1, estimated for the whole genome.
        \item $\ \proba [ \polyDel \given \divDel]$ is the proportion of DFE with a selection coefficient at the population-genetic scale ($\Spop$) lower than -1, estimated of SNPs observed with a selection coefficient at the phylogenetic scale ($\Sphy$) between -3 and -3.
        \item $\ \proba [ \polyDel \given \divStrongDel ]$ is the proportion of DFE with a selection coefficient at the population-genetic scale ($\Spop$) lower than -1, estimated of SNPs observed with a selection coefficient at the phylogenetic scale ($\Sphy$) lower than -3.
    \end{itemize}

    \subsubsection{Proportion of beneficial mutations ($\polyAdv$)}\label{subsec:proportion-beneficial-mutations}
    \captionof{figure}{~}
    \begin{minipage}{0.32\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/5bins-all/regression-MutSel/results.watterson.all_P-Ssup0.scatter.pdf}
    \end{minipage}
    \begin{minipage}{0.32\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/5bins-all/regression-MutSel/results.watterson.pos_P-Ssup0.scatter.pdf}
    \end{minipage}
    \begin{itemize}
        \item $ \proba [ \polyAdv ]$ is the proportion of DFE with a selection coefficient at the population-genetic scale ($\Spop$) greater than 1, estimated for the whole genome.
        \item $\proba [ \polyAdv \given \divAdv] $ is the proportion of DFE with a selection coefficient at the population-genetic scale ($\Spop$) greater than 1, estimated of SNPs observed with a selection coefficient at the phylogenetic scale ($\Sphy$) greater than 1.
    \end{itemize}

@@

%    \subsection{Selection efficacy correlation with diversity}\label{subsec:selection-efficacy-correlation-with-diversity}
%    The mean of the DFE $\SpopMean$ estimated by polyDFE is given by:
%    \begin{align*}
%        \SpopMean & = \int_{-\infty}^{+\infty} \Spop \phi \left( \Spop; \Spop_d , b, p_b, \Spop_b \right) \der \Spop \\
%        & = p_b \Spop_b + \left( 1 - p_b \right) \Spop_d
%    \end{align*}
%
%    To look further into the relation between diversity and selection efficacy, we studied the estimation of the mean $\SpopMean$ as a function of the mean $\SphyMean$ for a subset of SNPs, for which $\SphyMean$ is close to $0$.
%    Indeed, polyDFE is a nearly-neutral model and is not able to reliably compute selection coefficients for only strongly deleterious or only strongly beneficial SNPs.
%    Sliding windows of $5,000$ SNPs are taken with a mean $\SphyMean$ between $-0.5$ and $0.5$.
%    $\SpopMean$ is estimated for each group, with some groups that might have part of same SNPs that overlap.
%    The relationship between $\SpopMean$ and $\SphyMean$ can be described by a second-order polynomial, and the slope of this polynomial at $\SphyMean=0$ gives the ratio $\SpopMean/\SphyMean$ for each population.
%    \captionof{figure}{~}
%    \begin{center}
%        \includegraphics[width=\linewidth, page=1]{artworks/figure.slope-beta-S}
%    \end{center}
%
%    We observe that $\SpopMean/\SphyMean$, representing the efficacy of selection at the short versus long-term is positively correlated with synonymous diversity, a proxy of effective population size.
%    Altogether, populations with a large effective population size are under more efficient selection, confirming the prediction of the nearly-neutral theory of evolution that mutations are less likely to be considered neutral in populations with high diversity (see~\ref{subsec:proportion-nearly-neutral-mutations}, left scatter-plot), especially for mutations predicted already to be nearly-neutral at the phylogenetic scale (see~\ref{subsec:proportion-nearly-neutral-mutations}, middle and right scatter-plot).
%    As a consequence of an increased selection efficiency, mutations in populations with high diversity are mechanistically more likely to be deleterious (see~\ref{subsec:proportion-deleterious-mutations}, left scatter-plot), especially for mutations predicted already to be deleterious at the phylogenetic scale (see~\ref{subsec:proportion-deleterious-mutations}, middle and right scatter-plot).
%    A decrease in the proportion of neutral mutation did not translate into a mechanical increase in the proportion of beneficial mutations.
%    Instead, populations with higher diversity have a lower proportion of beneficial mutations, this effect is also more pronounced for mutations predicted to be beneficial at the phylogenetic scale.
%    This discrepancy can be explained by the fact that populations with larger diversity have already depleted the putatively beneficial mutations, and that only deleterious are remaining possible.
%    In other words, being on the top of a fixed fitness landscape leaves you with only paths going downhill.
%    Altogether, populations with increased diversity are showing signature of increased selection efficiency and are also less likely to generate positively selected mutations.
%   \newpage

    \subsection{Proportion of selected mutations}

    For each population, the proportion of beneficial ($p_+=\PpolyAdv$), nearly-neutral ($\PpolyNeutral$) and of deleterious ($\PpolyDel$) mutations estimated at the population-genetic scale is shown.
    Estimation is performed with SNPs with different class of selection at the phylogenetic scale: $\Sphyclass \in \{ \divStrongDel, \divDel, \divWeakDel, \divWeakAdv, \divAdv \}$
    Population are sorted for increased synonymous diversity.

>>>>

\newcommand{\UniDimArray}[1]{\bm{#1}}
\newcommand{\BiDimArray}[1]{\bm{#1}}
\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\Var}{\textrm{Var}}
\newcommand{\der}{\textrm{d}}
\newcommand{\e}{\textrm{e}}
\newcommand{\avg}[1]{\left< #1 \right>} % for average
\newcommand{\Ne}{N_{\textrm{e}}}
\newcommand{\proba}{\mathbb{P}}
\newcommand{\pfix}{\proba_{\textrm{fix}}}
\newcommand{\dn}{d_N}
\newcommand{\ds}{d_S}
\newcommand{\dnds}{\dn / \ds}
\newcommand{\Sphy}{S_{0}}
\newcommand{\Sphyclass}{\mathcal{C}}
\newcommand{\SphyMean}{\overline{\Sphy}}
\newcommand{\divDel}{\Sphy^{\text{d}}}
\newcommand{\divWeak}{\Sphy^{\text{n}}}
\newcommand{\divAdv}{\Sphy^{\text{b}}}
\newcommand{\PdivDel}{\proba \left[ \divDel \right]}
\newcommand{\PdivWeak}{\proba \left[ \divWeak \right]}
\newcommand{\PdivAdv}{\proba \left[ \divAdv \right]}
\newcommand{\given}{\mid}
\newcommand{\Spop}{S}
\newcommand{\SpopMean}{\overline{\Spop}}
\newcommand{\polyDel}{\Spop^{\text{d}}}
\newcommand{\polyNeutral}{\Spop^{\text{n}}}
\newcommand{\polyAdv}{\Spop^{\text{b}}}
\newcommand{\PpolyDel}{\proba \left[ \polyDel \right]}
\newcommand{\PpolyNeutral}{\proba \left[ \polyNeutral \right]}
\newcommand{\PpolyAdv}{\proba \left[ \polyAdv \right]}

\title{\textbf{Empirical evidence for beneficial mutations that are not adaptive innovations in mammals}}

\author{
    \large
    \textbf{T. {Latrille}$^{1}$, J. {Joseph}$^{2}$, D.~A. {Hartasánchez}$^{1}$, N. {Salamin}$^{1}$}\\
    \scriptsize $^{1}$Department of Computational Biology, Université de Lausanne, Lausanne, Switzerland\\
    \scriptsize $^{2}$Université de Lyon, Université Lyon 1, CNRS, VetAgro Sup, Laboratoire de Biométrie et Biologie Evolutive, UMR5558, Villeurbanne, France \\
    \texttt{\href{mailto:thibault.latrille@ens-lyon.org}{thibault.latrille@ens-lyon.org}} \\
}

@@

    \section{Mutation-selection codon models}\label{sec:site-specific-mutation-selection-codon-models}
    In BayesCode (\url{https://github.com/ThibaultLatrille/bayescode}), mutation-selection codon models are obtained by running \textit{mutselomega} for 2000 points of MCMC with the options:
    \begin{scriptsize}
        \begin{verbatim}
 mutselomega ---omegashift 0.0 --ncat 30 -a my_alignment.phy -t my_tree.newick -u 2000 my_genename
        \end{verbatim}
    \end{scriptsize}
    The collection of site-specific fitness profiles ($\UniDimArray{F^{(i)}}, \forall i$) are then obtained by running \textit{readmutselomega}, reading 1000 points of MCMC (first 1000 are considered as burn-in) with the options:
    \begin{scriptsize}
        \begin{verbatim}
 readmutselomega --every 1 --until 2000 --burnin 1000 --ss my_genename
        \end{verbatim}
    \end{scriptsize}
    The gene-specific $4 \times 4$ nucleotide mutation rate matrix ($\UniDimArray{\mu}$) is also obtained by running \textit{readmutselomega}, reading 1000 points of MCMC (first 1000 are considered as burn-in) with the options:
    \begin{scriptsize}
        \begin{verbatim}
 readmutselomega --every 1 --until 2000 --burnin 1000 --nuc my_genename
        \end{verbatim}
    \end{scriptsize}

@@

    \subsubsection{$\dnds$ and beneficial back-mutations}
    \begin{center}
        \captionof{table}{$\dnds$ of beneficial back-mutations.}
        \scriptsize
        \begin{longtable*}{|l|l|r|r|r|r|r|r|}
            \toprule
            Population &             Species & Tajima $\theta_{\pi}$  & $\mathbb{P}_{div}[ S_0^{b} ]$ & $\omega $ & $\omega ( S_0^{d} )$ & $\omega ( S_0^{n} )$ & $\omega ( S_0^{b} )$ \\
            \midrule
            \endfirsthead

            \toprule
            Population &             Species & Tajima $\theta_{\pi}$  & $\mathbb{P}_{div}[ S_0^{b} ]$ & $\omega $ & $\omega ( S_0^{d} )$ & $\omega ( S_0^{n} )$ & $\omega ( S_0^{b} )$ \\
            \midrule
            \endhead
            \midrule
            \multicolumn{8}{r}{{Continued on next page}} \\
            \midrule
            \endfoot

@@

            Equus c. &      Equus caballus &               $ 0.002$ &                      $ 0.118$ &  $ 0.147$ &             $ 0.075$ &             $ 0.823$ &             $ 1.203$ \\
            Iran &          Bos taurus &               $ 0.002$ &                      $ 0.123$ &  $ 0.131$ &             $ 0.074$ &             $ 0.627$ &             $ 1.206$ \\
            Uganda &          Bos taurus &               $ 0.005$ &                      $ 0.125$ &  $ 0.134$ &             $ 0.075$ &             $ 0.637$ &             $ 1.252$ \\
            Australia &        Capra hircus &               $ 0.003$ &                      $ 0.121$ &  $ 0.125$ &             $ 0.068$ &             $ 0.629$ &             $ 1.099$ \\
            France &        Capra hircus &               $ 0.004$ &                      $ 0.120$ &  $ 0.126$ &             $ 0.068$ &             $ 0.633$ &             $ 1.102$ \\
            Iran (C. aegagrus) &        Capra hircus &               $ 0.004$ &                      $ 0.120$ &  $ 0.125$ &             $ 0.068$ &             $ 0.630$ &             $ 1.100$ \\
            Iran &        Capra hircus &               $ 0.004$ &                      $ 0.121$ &  $ 0.125$ &             $ 0.067$ &             $ 0.630$ &             $ 1.108$ \\
            Italy &        Capra hircus &               $ 0.004$ &                      $ 0.120$ &  $ 0.126$ &             $ 0.068$ &             $ 0.634$ &             $ 1.102$ \\
            Morocco &        Capra hircus &               $ 0.004$ &                      $ 0.122$ &  $ 0.125$ &             $ 0.067$ &             $ 0.631$ &             $ 1.112$ \\
            Iran &          Ovis aries &               $ 0.007$ &                      $ 0.109$ &  $ 0.146$ &             $ 0.091$ &             $ 0.616$ &             $ 1.148$ \\
            Iran (O. orientalis) &          Ovis aries &               $ 0.009$ &                      $ 0.108$ &  $ 0.148$ &             $ 0.093$ &             $ 0.617$ &             $ 1.145$ \\
            Iran (O. vignei) &          Ovis aries &               $ 0.007$ &                      $ 0.109$ &  $ 0.146$ &             $ 0.091$ &             $ 0.617$ &             $ 1.134$ \\
            Various &          Ovis aries &               $ 0.008$ &                      $ 0.107$ &  $ 0.148$ &             $ 0.093$ &             $ 0.617$ &             $ 1.134$ \\
            Morocco &          Ovis aries &               $ 0.008$ &                      $ 0.108$ &  $ 0.148$ &             $ 0.093$ &             $ 0.619$ &             $ 1.149$ \\
            Barbados & Chlorocebus sabaeus &               $ 0.004$ &                      $ 0.122$ &  $ 0.133$ &             $ 0.071$ &             $ 0.696$ &             $ 1.397$ \\
            Central Afr. Rep. & Chlorocebus sabaeus &               $ 0.006$ &                      $ 0.124$ &  $ 0.134$ &             $ 0.071$ &             $ 0.701$ &             $ 1.431$ \\
            Ethiopia & Chlorocebus sabaeus &               $ 0.005$ &                      $ 0.124$ &  $ 0.134$ &             $ 0.071$ &             $ 0.699$ &             $ 1.422$ \\
            Gambia & Chlorocebus sabaeus &               $ 0.005$ &                      $ 0.123$ &  $ 0.134$ &             $ 0.071$ &             $ 0.703$ &             $ 1.411$ \\
            Kenya & Chlorocebus sabaeus &               $ 0.005$ &                      $ 0.123$ &  $ 0.134$ &             $ 0.071$ &             $ 0.699$ &             $ 1.419$ \\
            Nevis & Chlorocebus sabaeus &               $ 0.003$ &                      $ 0.123$ &  $ 0.133$ &             $ 0.071$ &             $ 0.699$ &             $ 1.407$ \\
            South Africa & Chlorocebus sabaeus &               $ 0.006$ &                      $ 0.125$ &  $ 0.134$ &             $ 0.070$ &             $ 0.704$ &             $ 1.447$ \\
            Saint Kitts & Chlorocebus sabaeus &               $ 0.004$ &                      $ 0.123$ &  $ 0.133$ &             $ 0.070$ &             $ 0.699$ &             $ 1.413$ \\
            Zambia & Chlorocebus sabaeus &               $ 0.006$ &                      $ 0.123$ &  $ 0.134$ &             $ 0.071$ &             $ 0.699$ &             $ 1.410$ \\
            African &        Homo sapiens &               $ 0.002$ &                      $ 0.099$ &  $ 0.192$ &             $ 0.118$ &             $ 0.865$ &             $ 1.609$ \\
            Admixed American &        Homo sapiens &               $ 0.002$ &                      $ 0.099$ &  $ 0.192$ &             $ 0.118$ &             $ 0.865$ &             $ 1.607$ \\
            East Asian &        Homo sapiens &               $ 0.002$ &                      $ 0.098$ &  $ 0.192$ &             $ 0.118$ &             $ 0.869$ &             $ 1.607$ \\
            European &        Homo sapiens &               $ 0.002$ &                      $ 0.098$ &  $ 0.192$ &             $ 0.118$ &             $ 0.865$ &             $ 1.598$ \\
            South Asian &        Homo sapiens &               $ 0.002$ &                      $ 0.099$ &  $ 0.192$ &             $ 0.118$ &             $ 0.867$ &             $ 1.610$ \\

@@

    \subsection{Correlation with diversity}
    For each population, the proportion of beneficial ($\PpolyAdv$), nearly-neutral ($\PpolyNeutral$) and of deleterious ($\PpolyDel$) mutations estimated at the population-genetic scale is shown as function of Watterson's diversity ($\Theta_w$).

    \subsubsection{Proportion of nearly-neutral mutations ($\polyNeutral$)}\label{subsec:proportion-nearly-neutral-mutations}
    \captionof{figure}{~}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC-masked/regression-MutSel/results.watterson.all_P-Sweak.scatter.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC-masked/regression-MutSel/results.watterson.weak_P-Sweak.scatter.pdf}
    \end{minipage}
    \begin{itemize}
        \item $\ \proba [ \polyNeutral ]$ is the proportion of DFE with a selection coefficient at the population-genetic scale ($\Spop$) between -1 and 1, estimated for the whole genome.
        \item $\ \proba [ \polyNeutral \given \divWeak ]$ is the proportion of DFE with a selection coefficient at the population-genetic scale ($\Spop$) between -1 and 1, estimated of SNPs observed with a selection coefficient at the phylogenetic scale ($\Sphy$) between -1 and 1.
    \end{itemize}

    We confirmed that higher synonymous diversity is typically accompanied by a smaller proportion of neutral mutations.
    This result is specially visible when we restricted the analysis to class of mutations that are supposedly nearly-neutral at the phylogenetic scale ($\divWeak$).
    This result suggests that populations with higher diversity (e.g.~\textit{Bos} or \textit{Ovis}) are more likely to discriminate whether mutations are beneficial or deleterious.
    Alternatively stated, mutations in populations with low diversity (e.g.~\textit{Homo}) are effectively nearly-neutral and behave as would a neutral mutation.
    This result is qualitatively in accordance with the nearly-neutral theory of evolution which argues that mutations are less efficiently selected for in small populations.

    \subsubsection{Proportion of deleterious mutations ($\polyDel$)}\label{subsec:proportion-deleterious-mutations}
    \captionof{figure}{~}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC-masked/regression-MutSel/results.watterson.all_P-Sneg.scatter.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC-masked/regression-MutSel/results.watterson.neg_P-Sneg.scatter.pdf}
    \end{minipage}
    \begin{itemize}
        \item $\ \proba [ \polyDel ]$ is the proportion of DFE with a selection coefficient at the population-genetic scale ($\Spop$) lower than -1, estimated for the whole genome.
        \item $\ \proba [ \polyDel \given \divDel]$ is the proportion of DFE with a selection coefficient at the population-genetic scale ($\Spop$) lower than -1, estimated of SNPs observed with a selection coefficient at the phylogenetic scale ($\Sphy$) lower than -1.
    \end{itemize}

    \subsubsection{Proportion of beneficial mutations ($\polyAdv$)}\label{subsec:proportion-beneficial-mutations}
    \captionof{figure}{~}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC-masked/regression-MutSel/results.watterson.all_P-Spos.scatter.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{../experiments/3bins-mC-masked/regression-MutSel/results.watterson.pos_P-Spos.scatter.pdf}
    \end{minipage}
    \begin{itemize}
        \item $ \proba [ \polyAdv ]$ is the proportion of DFE with a selection coefficient at the population-genetic scale ($\Spop$) greater than 1, estimated for the whole genome.
        \item $\proba [ \polyAdv \given \divAdv] $ is the proportion of DFE with a selection coefficient at the population-genetic scale ($\Spop$) greater than 1, estimated of SNPs observed with a selection coefficient at the phylogenetic scale ($\Sphy$) greater than 1.
    \end{itemize}

@@

    \section{Clinically related terms for non-synonymous mutations}

    \begin{center}
        \captionof{table}{Clinically related terms for $\Sphy < -1$}
        \begin{tabular}{|l|r|r|r|r|r|}
            \toprule
            SNP clinical ontology & $n_{\mathrm{Observed}}$ & $n_{\mathrm{Expected}}$ & Odds ratio & $p_{\mathrm{v}}$ & $p_{\mathrm{v-adjusted}}$ \\
            \midrule
            Benign                & 2969                    & $4043.0$                & $ 0.734$   & $ 1.000$             & $ 1.000~~$                    \\
            Likely benign         & 2994                    & $3399.8$                & $ 0.881$   & $ 0.999$             & $ 1.000~~$                    \\
            Risk factor           & 102                     & $ 118.2$                & $ 0.863$   & $ 0.798$             & $ 1.000~~$                    \\
            Likely pathogenic     & 221                     & $  68.5$                  & $ 3.226$   & $1.7\times 10^{-8}$  & $\bm{6.7\times 10^{-8}{^*}}$  \\
            Pathogenic            & 560                     & $ 193.6$                & $ 2.893$   & $4.2\times 10^{-17}$ & $\bm{2.1\times 10^{-16}{^*}}$ \\
            \bottomrule
        \end{tabular}
    \end{center}
    In humans, non-synonymous SNPs in the test group (deleterious mutations with $\Sphy < -1$) are contrasted to SNPs in the control group (nearly-neutral mutations with $-1 < \Sphy < 1$).
    For each clinical term, a 2x2 contingency tables is built by counting the number of SNPs based on their selection coefficient and their clinical terms (whether they have this specific term or not).
    Fisher's exact tests are then performed for these 2x2 contingency tables.
    $^*$ signify that the$p_{\mathrm{v}}^{\mathrm{adj}}$ corrected for multiple comparison (Holm–Bonferroni correction) is lower than the risk $\alpha=0.05$.
    SNPs predicted with $\Sphy < -1$ are statistically associated to clinical terms such as \textit{Likely Pathogenic} and \textit{Pathogenic}.

    \begin{center}
        \captionof{table}{Clinically related terms for $\Sphy > 1$}
        \begin{tabular}{|l|r|r|r|r|r|}
            \toprule
            SNP clinical ontology & $n_{\mathrm{Observed}}$ & $n_{\mathrm{Expected}}$ & Odds ratio & $p_{\mathrm{v}}$ & $p_{\mathrm{v-adjusted}}$ \\
            \midrule
            Benign                & 319                     & $ 261.7$                & $ 1.219$   & $ 0.002$         & $\bm{ 0.009{^*}}$         \\
            Likely benign         & 263                     & $ 222.7$                & $ 1.181$   & $ 0.012$         & $\bm{ 0.049{^*}}$         \\
            Risk factor           & 5                       & $ 7.847$                & $ 0.637$   & $ 0.879$         & $ 0.879~~$                \\
            Likely pathogenic     & 7                       & $ 4.552$                & $ 1.538$   & $ 0.227$         & $ 0.682~~$                \\
            Pathogenic            & 16                      & $  12.9$                  & $ 1.241$   & $ 0.268$         & $ 0.682~~$                \\
            \bottomrule
        \end{tabular}
    \end{center}
    In humans, non-synonymous SNPs in the test group (beneficial back-mutations with $\Sphy > 1$) are contrasted to SNPs in the control group (nearly-neutral mutations with $-1 < \Sphy < 1$).
    For each clinical term, a 2x2 contingency tables is built by counting the number of SNPs based on their selection coefficient and their clinical terms (whether they have this specific term or not).
    Fisher's exact tests are then performed for these 2x2 contingency tables.
    $^*$ signify that the$p_{\mathrm{v}}^{\mathrm{adj}}$ corrected for multiple comparison (Holm–Bonferroni correction) is lower than the risk $\alpha=0.05$.
    Beneficial back-mutations $\Sphy > 1$ are associated with clinical terms such as \textit{Benign} and \textit{Likely Benign}.

    \section{Distribution of fitness effects}
    \captionof{figure}{~}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{artworks/2bins.Yeast.experimentalDFE.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{artworks/2bins.AFR.histogramDFE.pdf}
    \end{minipage}
    Left panel: distribution of fitness effects experimentally determined in Yeast, data obtained in \textcite{shen_synonymous_2022}.
    Right panel:  distribution of fitness effects ($\Sphy$) as predicted with mutation-selection codon models, from reconstructed ancestral genome in \textit{Homo sapiens}.
    The DFE predicted with mutation-selection codon models has a negative mode, consistently with experimental DFEs, meaning that most mutations are effectively deleterious.
